---
title: "HBOI, SLR project, microarray differential expression data analysis""
author: Lisa Cohen (ljcohen@ucdavis.edu)
date: "`r Sys.Date()`"
output: html_document
---

```{r,echo=FALSE, message=FALSE, warning=FALSE}
   library("limma")
   library("gplots")
   library("arrayQualityMetrics")
   library("Biobase")
   library("ggplot2")
```

# Table of Contents:
1. Experiment Description
2. Data Analysis Procedure
3. Raw Data QC
4. Sample Names
5. PCA Results Figures
6. Results Files
7. Venn Diagram Results Figures
8. Comments
9. Contact
10. References

## 1. Experiment Description

This is a differential gene expression analysis from eleven custom Agilent microarray chips processed at HBOI on six dates: 12/5/2013, 5/24/2013, 5/30/2013, 6/11/2013, 6/5/2013, 6/6/2013. The experiment consists of hard coral samples, Montastraea cavernosa collected from 3 sites on 3 different dates. At least 4 samples at each site were collected from separate colonies for a total of 44 samples.

# 2. Data analysis procedure

Microarray chips were imaged at 635 nm using a high-resolution laser GenePix® 4200A microarray scanner (Axon, Molecular Devices, California, USA).  Data were extracted from images using GenePixPro 6.0 software as .gpr files.

Data were imported into the R statistical programming environment. The R-Bioconductor package, limma (Linear Models for Microarray Data, version 3.22.3, Ritchie et al. 2015; Smyth 2004) was used to assess the quality of the data and to analyze for differential expression.

```{r,echo=FALSE, message=FALSE, warning=FALSE}
options(width=600)
sample_info<-read.csv("targets_experiment_sample_info_revised_again.csv")
sample_info_matrix<-as.matrix(sample_info)
sample_info_matrix<-sample_info_matrix[,-1]
rownames(sample_info_matrix)<-sample_info$X
sample_info_dataframe<-as.data.frame(sample_info_matrix)
experimentData<-new("MIAME",name="Lisa Cohen, Joshua Voss, Sara Edge, Jeff Beal",
                    lab="Marine Functional Genomics, HBOI",
                    contact="ljcohen@ucdavis.edu",
                    title="St. Lucie Reef Seasonal Samples 2011",
                    abstract="ExpressionSet for SLR 2011 Microarray Data",
                    other=list(notes="Created from GenePix *.gpr files"))
pd<-new("AnnotatedDataFrame",data=sample_info_dataframe)
SLR_sum<-table(sample_info_dataframe$Site,sample_info_dataframe$Date)
SLR_sum
sample_info_dataframe
```


# 3. Raw Data QC:

Raw microarray data: 

```{r,echo=FALSE, message=FALSE, warning=FALSE}
   library("limma")
   library("gplots")
   library("arrayQualityMetrics")
   library("Biobase")
   library("ggplot2")
```

```{r,echo=FALSE, message=FALSE, warning=FALSE}
setwd("~/Documents/HBOI/SLR/SLR/SLR_2012_2013_rawdata_Rscriptsandanalyses")
#first chip name, array 4 image
chip1_4<-"63600_2013-05-24_250_array4.gpr"
#reads chip
chip1_4alldata<-readGAL(chip1_4,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
#takes only data from array 4
chip1_array4_alldata<-subset(chip1_4alldata,Array..==4)
chip1_array1_alldata<-subset(chip1_4alldata,Array..==1)
#first column is Name of feature
names<-data.frame(chip1_array1_alldata$Name)
colnames(names)<-"Name"
#in the future, you might want to know what Row and Column data come from
#for now, column and row are unnecessary because
#assume that all data will be in the same order
#and feature name will be the same for all arrays,  
#mydata<-data.frame(chip_data$Row,chip_data$Column,chip_data$Name)
#
#takes data Median feature intensity at 635 nm - Background intensity at 635 nm 
chip1_array4_data<-data.frame(chip1_array4_alldata$F635.Median...B635)
colnames(chip1_array4_data)<-"1_4"
all_data<-cbind(names,chip1_array4_data)
#chip 1 name, array3 image
chip1_3<-"63600_2013-05-24_300_array3.gpr"
chip1_3alldata<-readGAL(chip1_3,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip1_array3_alldata<-subset(chip1_3alldata,Array..==3)
chip1_array3_data<-data.frame(chip1_array3_alldata$F635.Median...B635)
colnames(chip1_array3_data)<-"1_3"
all_data<-cbind(all_data,chip1_array3_data)
chip1_2<-"63600_2013-05-24_370_arrays1,2.gpr"
chip1_2alldata<-readGAL(chip1_2,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip1_array2_alldata<-subset(chip1_2alldata,Array..==2)
chip1_array2_data<-data.frame(chip1_array2_alldata$F635.Median...B635)
colnames(chip1_array2_data)<-"1_2"
all_data<-cbind(all_data,chip1_array2_data)
chip1_1<-"63600_2013-05-24_300_array1.gpr"
chip1_1alldata<-readGAL(chip1_1,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip1_array1_alldata<-subset(chip1_1alldata,Array..==1)
chip1_array1_data<-data.frame(chip1_array1_alldata$F635.Median...B635)
colnames(chip1_array1_data)<-"1_1"
all_data<-cbind(all_data,chip1_array1_data)
chip2_4<-"63600_2013-12-05_0220_array4.gpr"
chip_alldata<-readGAL(chip2_4,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip2_array4_alldata<-subset(chip_alldata,Array..==4)
chip2_array4_data<-data.frame(chip2_array4_alldata$F635.Median...B635)
colnames(chip2_array4_data)<-"2_4"
all_data<-cbind(all_data,chip2_array4_data)
chip2_3<-"63600_2013-12-05_0220_array3.gpr"
chip2_3alldata<-readGAL(chip2_3,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip2_array3_alldata<-subset(chip2_3alldata,Array..==3)
chip2_array3_data<-data.frame(chip2_array3_alldata$F635.Median...B635)
colnames(chip2_array3_data)<-"2_3"
all_data<-cbind(all_data,chip2_array3_data)
chip2_2<-"63600_2013-12-05_0260_array2.gpr"
chip2_2alldata<-readGAL(chip2_2,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip2_array2_alldata<-subset(chip2_2alldata,Array..==2)
chip2_array2_data<-data.frame(chip2_array2_alldata$F635.Median...B635)
colnames(chip2_array2_data)<-"2_2"
all_data<-cbind(all_data,chip2_array2_data)
chip2_1<-"63600_2013-12-05_0300_arrays1,2.gpr"
chip2_1alldata<-readGAL(chip2_1,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip2_array1_alldata<-subset(chip2_1alldata,Array..==1)
chip2_array1_data<-data.frame(chip2_array1_alldata$F635.Median...B635)
colnames(chip2_array1_data)<-"2_1"
all_data<-cbind(all_data,chip2_array1_data)
chip3_4<-"63601_2013-06-06_0240_array4.gpr"
chip3_4alldata<-readGAL(chip3_4,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip3_array4alldata<-subset(chip3_4alldata,Array..==4)
chip3_array4data<-data.frame(chip3_array4alldata$F635.Median...B635)
colnames(chip3_array4data)<-"3_4"
all_data<-cbind(all_data,chip3_array4data)
chip3_arrays123<-"63601_2013-06-06_300_arrays1,2,3.gpr"
chip3_arrays123alldata<-readGAL(chip3_arrays123,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip3_array3_alldata<-subset(chip3_arrays123alldata,Array..==3)
chip3_array3data<-data.frame(chip3_array3_alldata$F635.Median...B635)
colnames(chip3_array3data)<-"3_3"
all_data<-cbind(all_data,chip3_array3data)
chip3_array2_alldata<-subset(chip3_arrays123alldata,Array..==2)
chip3_array2_data<-data.frame(chip3_array2_alldata$F635.Median...B635)
colnames(chip3_array2_data)<-"3_2"
all_data<-cbind(all_data,chip3_array2_data)
chip3_array1_alldata<-subset(chip3_arrays123alldata,Array..==1)
chip3_array1_data<-data.frame(chip3_array1_alldata$F635.Median...B635)
colnames(chip3_array1_data)<-"3_1"
all_data<-cbind(all_data,chip3_array1_data)
chip4_4<-"63605_2013-06-11_array4.gpr"
chip4_4alldata<-readGAL(chip4_4,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip4_array4alldata<-subset(chip4_4alldata,Array..==4)
chip4_array4data<-data.frame(chip4_array4alldata$F635.Median...B635)
colnames(chip4_array4data)<-"4_4"
all_data<-cbind(all_data,chip4_array4data)
chip4_3<-"63605_2013-06-11_array3.gpr"
chip4_array3alldata<-readGAL(chip4_3,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip4_array3_alldata<-subset(chip4_array3alldata,Array..==3)
chip4_array3data<-data.frame(chip4_array3_alldata$F635.Median...B635)
colnames(chip4_array3data)<-"4_3"
all_data<-cbind(all_data,chip4_array3data)
chip4_arrays12<-"63605_2013-06-11_300_arrays1,2.gpr"
chip4_arrays12alldata<-readGAL(chip4_arrays12,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip4_array2_alldata<-subset(chip4_arrays12alldata,Array..==2)
chip4_array2_data<-data.frame(chip4_array2_alldata$F635.Median...B635)
colnames(chip4_array2_data)<-"4_2"
all_data<-cbind(all_data,chip4_array2_data)
chip4_array1_alldata<-subset(chip4_arrays12alldata,Array..==1)
chip4_array1_data<-data.frame(chip4_array1_alldata$F635.Median...B635)
colnames(chip4_array1_data)<-"4_1"
all_data<-cbind(all_data,chip4_array1_data)
chip5_4<-"63606_2013-06-11_240_array4.gpr"
chip5_4alldata<-readGAL(chip5_4,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip5_array4alldata<-subset(chip5_4alldata,Array..==4)
chip5_array4data<-data.frame(chip5_array4alldata$F635.Median...B635)
colnames(chip5_array4data)<-"5_4"
all_data<-cbind(all_data,chip5_array4data)
chip5_3<-"63606_2013-06-11_0270_array3.gpr"
chip5_array3alldata<-readGAL(chip5_3,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip5_array3_alldata<-subset(chip5_array3alldata,Array..==3)
chip5_array3data<-data.frame(chip5_array3_alldata$F635.Median...B635)
colnames(chip5_array3data)<-"5_3"
all_data<-cbind(all_data,chip5_array3data)
chip5_arrays12<-"63606_2013-06-11_300_arrays1,2.gpr"
chip5_arrays12alldata<-readGAL(chip5_arrays12,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip5_array2_alldata<-subset(chip5_arrays12alldata,Array..==2)
chip5_array2_data<-data.frame(chip5_array2_alldata$F635.Median...B635)
colnames(chip5_array2_data)<-"5_2"
all_data<-cbind(all_data,chip5_array2_data)
chip5_array1_alldata<-subset(chip5_arrays12alldata,Array..==1)
chip5_array1_data<-data.frame(chip5_array1_alldata$F635.Median...B635)
colnames(chip5_array1_data)<-"5_1"
all_data<-cbind(all_data,chip5_array1_data)
chip6_4<-"63608_2013-12-05_0270_array4.gpr"
chip6_4alldata<-readGAL(chip6_4,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip6_array4alldata<-subset(chip6_4alldata,Array..==4)
chip6_array4data<-data.frame(chip6_array4alldata$F635.Median...B635)
colnames(chip6_array4data)<-"6_4"
all_data<-cbind(all_data,chip6_array4data)
chip6_3<-"63608_2013-12-05_0270_array3.gpr"
chip6_array3alldata<-readGAL(chip6_3,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip6_array3_alldata<-subset(chip6_array3alldata,Array..==3)
chip6_array3data<-data.frame(chip6_array3_alldata$F635.Median...B635)
colnames(chip6_array3data)<-"6_3"
all_data<-cbind(all_data,chip6_array3data)
chip6_arrays12<-"63608_2013-12-05_0280_arrays1,2.gpr"
chip6_arrays12alldata<-readGAL(chip6_arrays12,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip6_array2_alldata<-subset(chip6_arrays12alldata,Array..==2)
chip6_array2_data<-data.frame(chip6_array2_alldata$F635.Median...B635)
colnames(chip6_array2_data)<-"6_2"
all_data<-cbind(all_data,chip6_array2_data)
chip6_array1_alldata<-subset(chip6_arrays12alldata,Array..==1)
chip6_array1_data<-data.frame(chip6_array1_alldata$F635.Median...B635)
colnames(chip6_array1_data)<-"6_1"
all_data<-cbind(all_data,chip6_array1_data)
chip7_4<-"63710_2013-05-24_array4.gpr"
chip7_4alldata<-readGAL(chip7_4,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip7_array4alldata<-subset(chip7_4alldata,Array..==4)
chip7_array4data<-data.frame(chip7_array4alldata$F635.Median...B635)
colnames(chip7_array4data)<-"7_4"
all_data<-cbind(all_data,chip7_array4data)
chip7_3<-"63710_2013-05-24_300_array3.gpr"
chip7_array3alldata<-readGAL(chip7_3,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip7_array3_alldata<-subset(chip7_array3alldata,Array..==3)
chip7_array3data<-data.frame(chip7_array3_alldata$F635.Median...B635)
colnames(chip7_array3data)<-"7_3"
all_data<-cbind(all_data,chip7_array3data)
chip7_arrays12<-"63710_2013-05-24_350_arrays1,2.gpr"
chip7_arrays12alldata<-readGAL(chip7_arrays12,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip7_array2_alldata<-subset(chip7_arrays12alldata,Array..==2)
chip7_array2_data<-data.frame(chip7_array2_alldata$F635.Median...B635)
colnames(chip7_array2_data)<-"7_2"
all_data<-cbind(all_data,chip7_array2_data)
chip7_array1_alldata<-subset(chip7_arrays12alldata,Array..==1)
chip7_array1_data<-data.frame(chip7_array1_alldata$F635.Median...B635)
colnames(chip7_array1_data)<-"7_1"
all_data<-cbind(all_data,chip7_array1_data)
chip8_4<-"63768_2013-05-30_230_array4.gpr"
chip8_4alldata<-readGAL(chip8_4,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip8_array4alldata<-subset(chip8_4alldata,Array..==4)
chip8_array4data<-data.frame(chip8_array4alldata$F635.Median...B635)
colnames(chip8_array4data)<-"8_4"
all_data<-cbind(all_data,chip8_array4data)
chip8_3<-"63768_2013-05-30_250_array3.gpr"
chip8_array3alldata<-readGAL(chip8_3,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip8_array3_alldata<-subset(chip8_array3alldata,Array..==3)
chip8_array3data<-data.frame(chip8_array3_alldata$F635.Median...B635)
colnames(chip8_array3data)<-"8_3"
all_data<-cbind(all_data,chip8_array3data)
chip8_arrays12<-"63768_2013-05-30_300_arrays1,2.gpr"
chip8_arrays12alldata<-readGAL(chip8_arrays12,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip8_array2_alldata<-subset(chip8_arrays12alldata,Array..==2)
chip8_array2_data<-data.frame(chip8_array2_alldata$F635.Median...B635)
colnames(chip8_array2_data)<-"8_2"
all_data<-cbind(all_data,chip8_array2_data)
chip8_array1_alldata<-subset(chip8_arrays12alldata,Array..==1)
chip8_array1_data<-data.frame(chip8_array1_alldata$F635.Median...B635)
colnames(chip8_array1_data)<-"8_1"
all_data<-cbind(all_data,chip8_array1_data)
chip9_arrays34<-"63769_2013-05-30_230_arrays3,4.gpr"
chip9_arrays34alldata<-readGAL(chip9_arrays34,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip9_array4_alldata<-subset(chip9_arrays34alldata,Array..==4)
chip9_array4_data<-data.frame(chip9_array4_alldata$F635.Median...B635)
colnames(chip9_array4_data)<-"9_4"
all_data<-cbind(all_data,chip9_array4_data)
chip9_array3_alldata<-subset(chip9_arrays34alldata,Array..==3)
chip9_array3_data<-data.frame(chip9_array3_alldata$F635.Median...B635)
colnames(chip9_array3_data)<-"9_3"
all_data<-cbind(all_data,chip9_array3_data)
chip9_arrays12<-"63769_2013-05-30_300_arrays1,2.gpr"
chip9_arrays12alldata<-readGAL(chip9_arrays12,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip9_array2_alldata<-subset(chip9_arrays12alldata,Array..==2)
chip9_array2_data<-data.frame(chip9_array2_alldata$F635.Median...B635)
colnames(chip9_array2_data)<-"9_2"
all_data<-cbind(all_data,chip9_array2_data)
chip9_array1_alldata<-subset(chip9_arrays12alldata,Array..==1)
chip9_array1_data<-data.frame(chip9_array1_alldata$F635.Median...B635)
colnames(chip9_array1_data)<-"9_1"
all_data<-cbind(all_data,chip9_array1_data)
chip10_all<-"63770_2013-06-05_0220_all.gpr"
chip10_arrays1234alldata<-readGAL(chip10_all,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip10_array4_alldata<-subset(chip10_arrays1234alldata,Array..==4)
chip10_array4_data<-data.frame(chip10_array4_alldata$F635.Median...B635)
colnames(chip10_array4_data)<-"10_4"
all_data<-cbind(all_data,chip10_array4_data)
chip10_array3_alldata<-subset(chip10_arrays1234alldata,Array..==3)
chip10_array3_data<-data.frame(chip10_array3_alldata$F635.Median...B635)
colnames(chip10_array3_data)<-"10_3"
all_data<-cbind(all_data,chip10_array3_data)
chip10_array2_alldata<-subset(chip10_arrays1234alldata,Array..==2)
chip10_array2_data<-data.frame(chip10_array2_alldata$F635.Median...B635)
colnames(chip10_array2_data)<-"10_2"
all_data<-cbind(all_data,chip10_array2_data)
chip10_array1_alldata<-subset(chip10_arrays1234alldata,Array..==1)
chip10_array1_data<-data.frame(chip10_array1_alldata$F635.Median...B635)
colnames(chip10_array1_data)<-"10_1"
all_data<-cbind(all_data,chip10_array1_data)
chip11_4<-"63771_2013-06-06_220_array4.gpr"
chip11_4alldata<-readGAL(chip11_4,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip11_array4alldata<-subset(chip11_4alldata,Array..==4)
chip11_array4data<-data.frame(chip11_array4alldata$F635.Median...B635)
colnames(chip11_array4data)<-"11_4"
all_data<-cbind(all_data,chip11_array4data)
chip11_arrays123<-"63771_2013-06-06_0250_arrays1,2,3.gpr"
chip11_arrays123alldata<-readGAL(chip11_arrays123,path=NULL,header=TRUE,sep="\t",quote="\"",skip=NULL,as.is=TRUE)
chip11_array3_alldata<-subset(chip11_arrays123alldata,Array..==3)
chip11_array3data<-data.frame(chip11_array3_alldata$F635.Median...B635)
colnames(chip11_array3data)<-"11_3"
all_data<-cbind(all_data,chip11_array3data)
chip11_array2_alldata<-subset(chip11_arrays123alldata,Array..==2)
chip11_array2_data<-data.frame(chip11_array2_alldata$F635.Median...B635)
colnames(chip11_array2_data)<-"11_2"
all_data<-cbind(all_data,chip11_array2_data)
chip11_array1_alldata<-subset(chip11_arrays123alldata,Array..==1)
chip11_array1_data<-data.frame(chip11_array1_alldata$F635.Median...B635)
colnames(chip11_array1_data)<-"11_1"
all_data<-cbind(all_data,chip11_array1_data)
all_data_nonames<-all_data[,-1]
matrix_alldata<-as.matrix(all_data_nonames)
matrix_alldata[matrix_alldata<0]<-0
SLR_matrix_shift1<-matrix_alldata+1
nams<-all_data$Name
rownames(SLR_matrix_shift1)=make.names(nams,unique=TRUE)
head(SLR_matrix_shift1)
SLR_matrix_shift1_log2<-log2(SLR_matrix_shift1)
norm_SLR_matrix_shift1_log2<-normalizeBetweenArrays(SLR_matrix_shift1_log2)
SLR_matrix_shift1_log2_Loess<-normalizeCyclicLoess(SLR_matrix_shift1_log2)
```

# PCA results
## chip

```{r,echo=FALSE, message=FALSE, warning=FALSE}
library(lattice)
x<-SLR_matrix_shift1_log2_Loess
pca = prcomp(t(x))
names = colnames(x)
fac= factor(c("1","1","1","1","2","2","2","2","3","3","3","3","4","4","4","4","5","5","5","5","6","6","6","6","7","7","7","7","8","8","8","8",
              "9","9","9","9","10","10","10","10","11","11","11","11","12","12","12","12"))
colours = c("red","blue","green","yellow","orange","purple","black","brown","gray","lightblue","deeppink")
xyplot(
  PC2 ~ PC1, groups=fac, data=as.data.frame(pca$x), pch=16, cex=2,
  panel=function(x, y, ...) {
    panel.xyplot(x, y, ...);
    ltext(x=x, y=y, labels=names, pos=1, offset=0.8, cex=1)
  },
  aspect = "fill", col=colours,
  #main = draw.key(key = list(rect = list(col = list(col=colours), text = list(levels(fac)), rep = FALSE)))
)
```

## hyb date
```{r,echo=FALSE, message=FALSE, warning=FALSE}
fac= factor(c("5/24/13","5/24/13","5/24/13","5/24/13",
              "12/5/13","12/5/13","12/5/13","12/5/13",
              "6/6/13","6/6/13","6/6/13","6/6/13",
              "6/11/13","6/11/13","6/11/13","6/11/13",
              "6/11/13","6/11/13","6/11/13","6/11/13",
              "12/5/13","12/5/13","12/5/13","12/5/13",
              "5/24/13","5/24/13","5/24/13","5/24/13",
              "5/30/13","5/30/13","5/30/13","5/30/13",
              "5/30/13","5/30/13","5/30/13","5/30/13",
              "6/5/13","6/5/13","6/5/13","6/5/13",
              "6/6/13","6/6/13","6/6/13","6/6/13"))
names=fac
colours = c("red","blue","green","yellow","orange","purple","black","brown")
xyplot(
  PC2 ~ PC1, groups=fac, data=as.data.frame(pca$x), pch=16, cex=2,
  panel=function(x, y, ...) {
    panel.xyplot(x, y, ...);
    ltext(x=x, y=y, labels=names, pos=1, offset=0.8, cex=1)
  },
  aspect = "fill", col=colours,
  #main = draw.key(key = list(rect = list(col = list(col=colours), text = list(levels(fac)), rep = FALSE)))
)
```
## site
```{r,echo=FALSE, message=FALSE, warning=FALSE}
fac= factor(c("South","Ledge","South","Ledge",
              "Central","South","Ledge","Ledge",
              "Ledge","South","Central","Ledge",
              "Central","South","Ledge","South",
              "South","South","Ledge","South",
              "Ledge","South","Central","Ledge",
              "Central","Central","South","Ledge",
              "South","Ledge","South","South",
              "Ledge","Central","South","Ledge",
              "Ledge","South","Central","Ledge",
              "South","South","Ledge","South"))
names=fac
colours = c("red","blue","green")
xyplot(
  PC2 ~ PC1, groups=fac, data=as.data.frame(pca$x), pch=16, cex=2,
  panel=function(x, y, ...) {
    panel.xyplot(x, y, ...);
    ltext(x=x, y=y, labels=names, pos=1, offset=0.8, cex=1)
  },
  aspect = "fill", col=colours,
  #main = draw.key(key = list(rect = list(col = list(col=colours), text = list(levels(fac)), rep = FALSE)))
)
```

## date
```{r,echo=FALSE, message=FALSE, warning=FALSE}
fac= factor(c("Oct2012","June2012","June2012","JanFeb2013",
              "JanFeb2013","Oct2012","JanFeb2013","Oct2012",
              "Oct2012","Oct2012","JanFeb2013","Oct2012",
              "June2012","June2012","June2012","June2012",
              "JanFeb2013","JanFeb2013","JanFeb2013","JanFeb2013",
              "June2012","June2012","JanFeb2013","JanFeb2013",
              "June2012","June2012","June2012","June2012",
              "JanFeb2013","JanFeb2013","JanFeb2013","JanFeb2013",
              "Oct2012","JanFeb2013","Oct2012","Oct2012",
              "Oct2012","June2012","June2012","June2012",
              "JanFeb2013","JanFeb2013","JanFeb2013","JanFeb2013"))
names=fac
colours = c("red","blue","green")
xyplot(
  PC2 ~ PC1, groups=fac, data=as.data.frame(pca$x), pch=16, cex=2,
  panel=function(x, y, ...) {
    panel.xyplot(x, y, ...);
    ltext(x=x, y=y, labels=names, pos=1, offset=0.8, cex=1)
  },
  aspect = "fill", col=colours,
  #main = draw.key(key = list(rect = list(col = list(col=colours), text = list(levels(fac)), rep = FALSE)))
)
```

## var
```{r,echo=FALSE, message=FALSE, warning=FALSE}
fac= factor(c("South_Oct2012","Ledge_June2012","South_June2012","Ledge_JanFeb2013",
              "Central_JanFeb2013","South_Oct2012","Ledge_JanFeb2013","Ledge_Oct2012",
              "Ledge_Oct2012","South_Oct2012","Central_JanFeb2013","Ledge_Oct2012",
              "Central_June2012","South_June2012","Ledge_June2012","South_June2012",
              "South_JanFeb2013","South_JanFeb2013","Ledge_JanFeb2013","South_JanFeb2013",
              "Ledge_June2012","South_June2012","Central_JanFeb2013","Ledge_JanFeb2013",
              "Central_June2012","Central_June2012","South_June2012","Ledge_June2012",
              "South_JanFeb2013","Ledge_JanFeb2013","South_JanFeb2013","South_JanFeb2013",
              "Ledge_Oct2012","Central_JanFeb2013","South_Oct2012","Ledge_Oct2012",
              "Ledge_Oct2012","South_June2012","Central_June2012","Ledge_June2012",
              "South_JanFeb2013","South_JanFeb2013","Ledge_JanFeb2013","South_JanFeb2013"))
names=fac
colours = c("red","blue","green","yellow","deeppink","black","lightblue","purple")
xyplot(
  PC2 ~ PC1, groups=fac, data=as.data.frame(pca$x), pch=16, cex=2,
  panel=function(x, y, ...) {
    panel.xyplot(x, y, ...);
    ltext(x=x, y=y, labels=names, pos=1, offset=0.8, cex=1)
  },
  aspect = "fill", col=colours,
  #main = draw.key(key = list(rect = list(col = list(col=colours), text = list(levels(fac)), rep = FALSE)))
)
```

```{r,echo=FALSE, message=FALSE, warning=FALSE}

```

# 9. References

Izarry RA, Hobbs B, Collin F, Beazer-Barclay YD, Antonellis KJ, Scherf U, Speed TP. (2003). Exploration, normalization, and summaries of high density oligonucleotide array probe level data. Biostatistics. 2003 Apr;4(2):249-64. http://www.ncbi.nlm.nih.gov/pubmed/12925520

Ritchie, ME, Phipson, B, Wu, D, Hu, Y, Law, CW, Shi, W, and Smyth, GK (2015). limma powers differential expression analyses for RNA-sequencing and microarray studies. Nucleic Acids Research 43(7), e47. http://www.ncbi.nlm.nih.gov/pubmed/25605792

Smyth, G. K. (2004). Linear models and empirical Bayes methods for assessing differential expression in microarray experiments. Statistical Applications in Genetics and Molecular Biology, Vol. 3, No. 1, Article 3. http://www.ncbi.nlm.nih.gov/pubmed/16646809 http://www.statsci.org/smyth/pubs/ebayes.pdf

R-Bioconductor: http://www.bioconductor.org/

limma: http://www.bioconductor.org/packages/release/bioc/vignettes/limma/inst/doc/usersguide.pdf