---
title: "Dasen lab, Pbx-mutant RNAseq", brachial/thoracic overlap
author: "Lisa Cohen"
output: html_document
---

.Rmd script used to generate this page here:
https://github.com/ljcohen/ljcohen.github.io/blob/master/analyses/Dasen/scripts/Dasen_RNAseq_report_brachial_thoracic_overlap.Rmd


Filenames for brachial-level transcripts are here:
```{r,echo=FALSE, message=FALSE, warning=FALSE}
library(DESeq2)
library("GenomicFeatures")
library("genefilter")
library(gplots)
library(RColorBrewer)
library(biomaRt)
library("genefilter")
library("lattice")
setwd("~/Documents/NYUMC/Dasen/brachial/htseq_counts")
mypath<-"~/Documents/NYUMC/Dasen/brachial/htseq_counts"
filenames<-list.files(path=mypath, pattern= "_counts.txt", full.names=FALSE)
datalist <-lapply(filenames, function(x){read.table(x,header=FALSE, sep="\t")})
for (i in 1:length(filenames))
{
  colnames(datalist[[i]])<-c("ID",filenames[[i]])
}
mergeddata <- Reduce(function(x,y) {merge(x,y, by="ID")}, datalist)
new_data_merge<-mergeddata[-1:-5,]
#write.csv(new_data_merge,file="Dasen_thoracic_count_data_Ensembl.csv")
rown<-new_data_merge$ID
rownames(new_data_merge)<-rown
new_data_merge<-new_data_merge[,-1]
data<-new_data_merge
colnames(data)
col.names<-c("BR-A-Control","BR-A-Mutant","BR-B-Control","BR-B-Mutant","BR-C-Control","BR-C-Mutant")
colnames(data)<-col.names
thoracic_data<-data
ExpDesign <- data.frame(row.names=colnames(thoracic_data), condition = c("Control","Mutant","Control","Mutant","Control","Mutant"))
cds<-DESeqDataSetFromMatrix(countData=data, colData=ExpDesign,design=~condition)
cds$condition <- relevel(cds$condition, "Control")
cds<-DESeq(cds, betaPrior=FALSE)
cds_brachial<-cds
```

Filenames for thoracic-level transcripts are here:
```{r,echo=FALSE, message=FALSE, warning=FALSE}
setwd("~/Documents/NYUMC/Dasen/thoracic/htseq_counts")
mypath<-"~/Documents/NYUMC/Dasen/thoracic/htseq_counts"
filenames<-list.files(path=mypath, pattern= "_counts.txt", full.names=FALSE)
datalist <-lapply(filenames, function(x){read.table(x,header=FALSE, sep="\t")})
for (i in 1:length(filenames))
{
  colnames(datalist[[i]])<-c("ID",filenames[[i]])
}
mergeddata <- Reduce(function(x,y) {merge(x,y, by="ID")}, datalist)
new_data_merge<-mergeddata[-1:-5,]
#write.csv(new_data_merge,file="Dasen_thoracic_count_data_Ensembl.csv")
rownames(new_data_merge)<-new_data_merge$ID
new_data_merge<-new_data_merge[,-1]
data<-new_data_merge
colnames(data)
col.names<-c("TH-A-Control","TH-A-Mutant","TH-B-Control","TH-B-Mutant","TH-C-Control","TH-C-Mutant")
colnames(data)<-col.names
brachial_data<-data
ExpDesign <- data.frame(row.names=colnames(brachial_data), condition = c("Control","Mutant","Control","Mutant","Control","Mutant"))
cds<-DESeqDataSetFromMatrix(countData=data, colData=ExpDesign,design=~condition)
cds$condition <- relevel(cds$condition, "Control")
cds<-DESeq(cds,betaPrior=FALSE)
cds_thoracic<-cds
```

The size of the table with all brachial transcripts is: 

```{r,echo=FALSE, message=FALSE, warning=FALSE}
# get norm counts
norm_counts<-counts(cds_brachial,normalized=TRUE)
norm_counts_data<-as.data.frame(norm_counts)
ensembl_id<-rownames(norm_counts)
norm_counts_data<-cbind(ensembl_id,norm_counts_data)
filtered_norm_counts<-norm_counts_data[!rowSums(norm_counts_data[,2:7]==0)>=1, ]
# get gene name from Ensembl gene ID
ensembl=useMart("ensembl")
ensembl = useDataset("mmusculus_gene_ensembl",mart=ensembl)
data_table<-filtered_norm_counts

query<-getBM(attributes=c('ensembl_gene_id','external_gene_name','gene_biotype'), filters = 'ensembl_gene_id', values = ensembl_id, mart=ensembl)
col.names<-c("ensembl_id","external_gene_id","gene_biotype")
colnames(query)<-col.names
merge_biomart_res_counts <- merge(data_table,query,by="ensembl_id")
temp_data_merged_counts<-merge_biomart_res_counts

##
res<-results(cds_brachial,contrast=c("condition","Mutant","Control"))
res_ordered<-res[order(res$padj),]
ensembl_id<-rownames(res_ordered)
res_ordered<-as.data.frame(res_ordered)
res_ordered<-cbind(res_ordered,ensembl_id)
merge_biomart_res_counts <- merge(temp_data_merged_counts,res_ordered,by="ensembl_id")
merge_biomart_res_all<-subset(merge_biomart_res_counts,merge_biomart_res_counts$padj!="NA")
merge_biomart_res_all<-merge_biomart_res_all[order(merge_biomart_res_all$padj),]
dim(merge_biomart_res_all)



```

The size of the brachial table with only significant transcripts, padj<0.05 is:
```{r,echo=FALSE, message=FALSE, warning=FALSE}
res_merged_cutoff<-subset(merge_biomart_res_all,merge_biomart_res_all$padj<0.05)
dim(res_merged_cutoff)
res1_filtered_padj<-res_merged_cutoff
```

The size of the table with all thoracic transcripts is: 

```{r,echo=FALSE, message=FALSE, warning=FALSE}
# get norm counts
norm_counts<-counts(cds_thoracic,normalized=TRUE)
norm_counts_data<-as.data.frame(norm_counts)
ensembl_id<-rownames(norm_counts)
norm_counts_data<-cbind(ensembl_id,norm_counts_data)
filtered_norm_counts<-norm_counts_data[!rowSums(norm_counts_data[,2:7]==0)>=1, ]
# get gene name from Ensembl gene ID
ensembl=useMart("ensembl")
ensembl = useDataset("mmusculus_gene_ensembl",mart=ensembl)
data_table<-filtered_norm_counts

query<-getBM(attributes=c('ensembl_gene_id','external_gene_name','gene_biotype'), filters = 'ensembl_gene_id', values = ensembl_id, mart=ensembl)
col.names<-c("ensembl_id","external_gene_id","gene_biotype")
colnames(query)<-col.names
merge_biomart_res_counts <- merge(data_table,query,by="ensembl_id")
temp_data_merged_counts<-merge_biomart_res_counts

##
res<-results(cds_thoracic,contrast=c("condition","Mutant","Control"))
res_ordered<-res[order(res$padj),]
ensembl_id<-rownames(res_ordered)
res_ordered<-as.data.frame(res_ordered)
res_ordered<-cbind(res_ordered,ensembl_id)
merge_biomart_res_counts <- merge(temp_data_merged_counts,res_ordered,by="ensembl_id")
merge_biomart_res_all<-subset(merge_biomart_res_counts,merge_biomart_res_counts$padj!="NA")
merge_biomart_res_all<-merge_biomart_res_all[order(merge_biomart_res_all$padj),]
dim(merge_biomart_res_all)

```

The size of the thoracic table with only significant transcripts, padj<0.05 is:
```{r,echo=FALSE, message=FALSE, warning=FALSE}
res_merged_cutoff<-subset(merge_biomart_res_all,merge_biomart_res_all$padj<0.05)
dim(res_merged_cutoff)
res2_filtered_padj<-res_merged_cutoff
```


# Venn Diagram

```{r,echo=FALSE, message=FALSE, warning=FALSE}
source('~/Documents/scripts/overLapper_original.R')
# brachial
m<-res1_filtered_padj$external_gene_id
# thoracic
n<-res2_filtered_padj$external_gene_id
setlist <- list(Brachial_v_Control=as.vector(m),Thoracic_v_Control=as.vector(n))
OLlist <- overLapper(setlist=setlist, sep="", type="vennsets")
counts <- sapply(OLlist$Venn_List, length)
vennPlot(counts=counts)
```

```{r,echo=FALSE, message=FALSE, warning=FALSE}
# extract intersections for up:
Brachial_Thoracic<-OLlist$Venn_List$Brachial_v_ControlThoracic_v_Control
Brachial_Overlap<-res1_filtered_padj[res1_filtered_padj$external_gene_id %in% Brachial_Thoracic,]
Thoracic_Overlap<-res2_filtered_padj[res2_filtered_padj$external_gene_id %in% Brachial_Thoracic,]
Brachial_Thoracic_results<-merge(Brachial_Overlap,Thoracic_Overlap,by="ensembl_id")
```

# Heatmap

```{r,echo=FALSE, message=FALSE, warning=FALSE}

d<-as.matrix(Brachial_Thoracic_results[,c(2:7,16:21)])
rownames(d) <- Brachial_Thoracic_results[,8]
d<-na.omit(d)
d<-d[,c(1,3,5,2,4,6,7,9,11,8,10,12)]
colnames(d)<-c("BR-A-Control","BR-B-Control","BR-C-Control","BR-A-Mutant","BR-B-Mutant","BR-C-Mutant","TH-A-Control","TH-B-Control","TH-C-Control","TH-A-Mutant","TH-B-Mutant","TH-C-Mutant")
hr <- hclust(as.dist(1-cor(t(d), method="pearson")), method="complete")
mycl <- cutree(hr, h=max(hr$height/1.5))
clusterCols <- rainbow(length(unique(mycl)))
myClusterSideBar <- clusterCols[mycl]
myheatcol <- greenred(75)
tiff("Overlap_heatmap.tiff", width = 1000,height = 1000,units="px",res = NA,pointsize=12)
heatmap.2(d, 
          Rowv=as.dendrogram(hr),
          cexRow=1,cexCol=0.8,srtCol= 90,
          adjCol = c(NA,0),offsetCol=2.5,offsetRow=0.1,
          Colv=NA, dendrogram="row", 
          scale="row", col=myheatcol, 
          density.info="none", 
          trace="none")
dev.off()
###

```

File:

```{r}
write.csv(Brachial_Thoracic_results,"Dasen_Brachial_Thoracic_merged_padj0.05.csv")
```

# Contact

Lisa Cohen, PhD student in MCIP, UC Davis     
Phone: 321-427-9335       
Email: ljcohen@ucdavis.edu

### Sequencing and original bioinformatics analysis by:

NYU Langone Medical Center   
Bioinformatics Core, Genome Technology Center, OCS   
Email: Genomics@nyumc.org         
Phone: 646-501-2834   
http://ocs.med.nyu.edu/bioinformatics-core  
http://ocs.med.nyu.edu/genome-technology-center   

Dr. Steven Shen, Associate Professor      
Email: ShuQuan.Shen@nyumc.org  
Phone: 212-263-2234           
Genome Technology Center, Office of Collaborative Sciences
and Center for Health Informatics and Bioinformatics
NYU School of Medicine,
227 E30th Street, Room 748, 
New York, NY 10016



# References

M. I. Love, W. Huber, S. Anders: Moderated estimation of fold change and dispersion for RNA-seq data with DESeq2.
Genome Biology 2014, 15:550. http://dx.doi.org/10.1186/s13059-014-0550-8

R-Bioconductor: http://www.bioconductor.org/

DESeq2: http://www.bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.pdf